Load libraries
```{r}
library(cowplot)
library(tidyverse)
```

Load data
```{r}
#Generate palette
ntwk.palette = c(rcartocolor::carto_pal(7, "Tropic")[2], "gray85", rcartocolor::carto_pal(7, "Tropic")[7])

#Load indicator species results
res.trt <- readRDS("../Data/votu_indicator.RDS") %>% 
  rename("Genome" = "OTU_ID")

#Filter significant vOTUs
sig.otus <- res.trt %>% 
  filter(p.value < 0.05) %>% 
  mutate(Trait = Response)

#Load safe network layout
nodes <- readRDS("../Data/safe_ind_whole_nodes.RDS")
edges <- readRDS("../Data/safe_ind_whole_edges.RDS")

#Load neighborhood enrichment results
enrichment <- readRDS("../Data/safe_ind_enrichment.RDS")

#Load neigbhorhood membership data frame
neighborhoods.tidy <- readRDS("../Data/safe_ind_neighborhoods.RDS")
```

Generate palette values for the enrichment plots
```{r}
enrichment.pal.trt <- ntwk.palette[c(2,7)]
names(enrichment.pal.trt) <- unique(sig.otus$Trait)

enr.color.list.trt <- list()

for(enr in sort(unique(sig.otus$Trait))){
  color = enrichment.pal.trt[[enr]]
  enr.color.list.trt[[enr]] <- data.frame(EnrichmentBin = 1:5,
                                      Color = colorRampPalette(c("white", color))(6)[2:6])
}

enr.color.trt <- plyr::ldply(enr.color.list.trt, function(x) x) %>% 
  dplyr::rename("Trait" = ".id") %>% 
  arrange(Trait, EnrichmentBin)
```

Plot the distribution of indicator species across the network and the corresponding neighborhood enrichments
```{r}
#Dataframe to plot outside circle 
circle <- data.frame(x0 = 0, y0 = 0, r = 0.5)

#Annotate network data frame with indicator species results
dao.tmp.trt <- nodes %>% 
  left_join(sig.otus, by = "Genome") %>% 
  mutate(Trait = ifelse(is.na(Trait), "NS", as.character(Trait))) %>% 
  mutate(Trait = fct_recode(Trait,
                            "Enriched in T2-50" = "Dry",
                            "Depleted in T2-50" = "Wet",
                            "P ≥ 0.5" = "NS"))

#Plot distribution of indicator species across the network
dao.trt.p <- dao.tmp.trt %>% 
  ggplot() +
  geom_line(data = edges, aes(x, y, group = Pair), alpha = 0.1, size = 0.1) +
  geom_point(data = filter(dao.tmp.trt, Trait == "P ≥ 0.5"), aes(x, y, color = Trait), alpha = 0.8, size = 0.5, shape = 16) +
  geom_point(data = filter(dao.tmp.trt, Trait != "P ≥ 0.5"), aes(x, y, color = Trait), alpha = 0.8, size = 1) +
  ggforce::geom_circle(data = circle, aes(x0 = 0.5, y0 = 0.5, r = 0.55), color = "gray25", linetype = 2) +
  scale_color_manual(values = ntwk.palette[c(3,1,2)], limits = c("Enriched in T2-50","Depleted in T2-50", "P ≥ 0.5")) +
  guides(color = guide_legend(title.position = "left", title.hjust = 1, nrow = 3)) +
  theme(text = element_text(size = 12),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom")

dao.trt.p

#Annotate network data frame with neighborhood enrichment results
enr.tmp.trt <- nodes %>%
  inner_join(enrichment, by = c("Genome" = "Neighborhood")) %>%
  arrange(EnrichmentScore) %>%
  filter(padj < 0.05)

#Plot neighbhorhood enrichment scores across the network
enr.trt.p <- enr.tmp.trt %>% 
  arrange(EnrichmentScore) %>% 
  mutate(Trait = fct_recode(Trait,
                            "Enriched in T2-50" = "Dry",
                            "Depleted in T2-50" = "Wet")) %>% 
  mutate(Trait = fct_relevel(Trait,
                             "Enriched in April 50%")) %>% 
  arrange(logpval) %>% 
  ggplot() +
  geom_point(aes(x,y, color = logpval), alpha = 1, size = 1) +
  scale_color_distiller(name = "Trait overrepresentation\n-log10(P-val)",
                        palette = "Blues", direction = 1) +
  guides(color = guide_colorbar(title.position = "left", title.hjust = 1)) +
  ggforce::geom_circle(data = circle, aes(x0 = 0.5, y0 = 0.5, r = 0.55), color = "gray25", linetype = 2) +
  facet_grid(. ~ Trait) +
  theme(text = element_text(size = 12),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        strip.background =element_rect(fill="gray25"),
        strip.text = element_text(colour = "white")) 

enr.trt.p

#Generate top panel 
top <- cowplot::plot_grid(dao.trt.p, enr.trt.p, rel_widths = c(1,2), align = "h", axis = "bt", labels = c("a", "b"), label_size = 15)
top
```

Generate dataframes summarizing indicator species results across relevant neighborhoods
```{r}
#Get all 
sig.nbhd <- enrichment %>% 
  filter(padj < 0.05) %>% 
  filter(Trait == "Dry")

#Generate a dataframe summarizing the number of neighborhood members identified as indicator species
response.summary <- neighborhoods.tidy %>% 
  filter(Neighborhood %in% sig.nbhd$Neighborhood) %>% 
  inner_join(res.trt, by = c("Members" = "Genome")) %>% 
  group_by(Neighborhood, Response) %>% 
  summarise(Total = n()) %>% 
  group_by(Neighborhood) %>% 
  mutate(Fraction = Total/sum(Total)) %>% 
  ungroup() %>% 
  mutate(Response = fct_relevel(Response,
                                c("Wet", "NS", "Dry"))) %>% 
  inner_join(select(enr.tmp.trt, Genome, logpval), by = c("Neighborhood" = "Genome")) %>% 
  mutate(Set = "Neighborhood")

#Rank neighborhoods by the fraction of members that were enriched in T2-50 samples
response.rank <- response.summary %>% 
  filter(Response == "Dry") %>% 
  mutate(Rank = rank(Fraction, ties.method = "first")) %>% 
  select(Neighborhood, Rank)

#Generate a dataframe summarizing the number of vOTU identified as indicator species across the whole network
response.whole <- nodes %>% 
  inner_join(res.trt, by = "Genome") %>% 
  group_by(Response) %>% 
  summarise(Total = n()) %>% 
  ungroup() %>% 
  mutate(Fraction = Total/sum(Total)) %>% 
  ungroup() %>% 
  mutate(Response = fct_relevel(Response,
                                c("Wet", "NS", "Dry"))) %>% 
  mutate(Neighborhood = "Whole",
         logpval = NA) %>% 
  select(Neighborhood, Response, Total, Fraction, logpval) %>% 
  mutate(Set = "Total network")
```


Bind response data frames 
```{r}
rbind(response.summary, response.whole) %>%
  mutate(Set = fct_relevel(Set, "Total network")) %>% 
  group_by(Set,Neighborhood) %>% 
  summarise(Total2 = sum(Total)) %>% 
  mutate(Total2 = ifelse(Set == "Total network", 0, Total2)) %>% 
  arrange(Total2)
```

Plot
```{r}
#Barplot displaying the size of each significant neighborhood
size.p <- rbind(response.summary, response.whole) %>%
  mutate(Set = fct_relevel(Set, "Total network")) %>% 
  group_by(Set,Neighborhood) %>% 
  summarise(Total2 = sum(Total)) %>% 
  mutate(Total2 = ifelse(Set == "Total network", 0, Total2)) %>% 
  left_join(response.rank, by = "Neighborhood") %>% 
  ggplot(aes(reorder(Neighborhood, -Rank), Total2)) +
  geom_bar(stat = "identity") +
  xlab("Network\nneighborhood") +
  ylab("Neighborhood size\n(# of nodes)") +
  facet_grid(. ~ Set, scales = "free", space = "free") +
  theme_minimal() +
  theme(text = element_text(size = 12),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_blank()) 

#Barplot displaying the composition of each neighborhood
response.p <- rbind(response.summary, response.whole) %>%
  mutate(Set = fct_relevel(Set, "Total network")) %>% 
  left_join(response.rank, by = "Neighborhood") %>% 
  ggplot() +
  geom_bar(aes(reorder(Neighborhood, -Rank), Fraction, fill = Response), stat = "identity", position = "fill") +
  geom_point(aes(reorder(Neighborhood, -Rank), -0.05, color = logpval)) +
   scale_color_distiller(palette = "Blues", direction = 1) +
  scale_fill_manual(name = "Enriched in April 50%\nDepleted in April 50%\nOther",
                    values = ntwk.palette,
                    limits = c("Wet", "NS", "Dry"),
                    labels = c("Depleted", "P ≥ 0.05", "Enriched")) +
  guides(color = FALSE) +
  guides(fill = guide_legend(title.position = "left", title.hjust = 1, nrow = 3)) +
  xlab("Network\nneighborhood") +
  ylab("Fraction of nodes") +
  facet_grid(. ~ Set, scales = "free", space = "free") +
  theme_minimal() +
  theme(text = element_text(size = 12),
        axis.text.x = element_blank(), strip.text = element_blank(),
        legend.position = "none") 

response.p

bottom <- plot_grid(size.p, response.p, align = "v", axis = "lr", nrow = 2, rel_heights = c(6,11))
```
Put it all togehter
740:830
```{r}
plot_grid(top, bottom, nrow = 2, labels = c(NA, "c"), label_size = 15, rel_heights = c(4,5))
```

Get some stats for the paper
```{r}
neighborhoods.tidy %>% 
  group_by(Neighborhood) %>% 
  count() %>%
  filter(n > 10) %>% 
  .$n %>% 
  median()

response.summary %>% 
  filter(Response == "Dry") %>% 
  arrange(Fraction)
```

